A Plot is Worth a Thousand Tests: Assessing Residual Diagnostics with the Lineup Protocol

### 2023-03-24 paper_comparison/paper_comparison/paper_comparison.tex
modify the x_dist text

\title{A plot is worth a thousand tests: assessing residual diagnostics
with visual inference}

@@

Figure \ref{fig:different-x-dist-poly-power} illustrates the change of
power of visual tests for different fitted distributions used in the
non-linearity model and the heteroskedasticity model. Note that we have
collected more evaluations on lineup with uniform fitted value
distribution, and the number of evaluations on a lineup will have an
impact on the power of the visual test. To have a fair comparison, the
five evaluations used in the power estimation are randomly sampled from
the total eleven evaluations for each lineup with uniform fitted value
distribution. The sampling process has been repeated 500 times to give
an indication of the variation of the power.

We do not observe significant difference between the power curve of the
uniform distribution and the power curve of the discrete uniform
distribution, but the power curves of normal and lognormal distributions
are slightly lower. Same can be observed for heteroskedasticity
patterns, where the gap between the power curve of the discrete uniform
and the power curve of the normal distribution is even larger. This
indicates that residual departures displayed with evenly distributed
fitted values are visually more attractive. Besides, the power of
detecting non-linearity or heteroskedasticity patterns from a lineup
with lognormal fitted value distribution is the lowest among four
distributions. This suggests the skewness of the fitted value
distribution plays a role in the degree of presence of visual features,
and a symmetric fitted value distribution can better reveal the
underlying shape.

@@

\caption{Comparison of power on lineups with different fitted value distributions for visual tests. All the power curves are estimated using five evaluations on each lineup. For lineups with uniform fitted value distribution, the five evaluations are randomly sampled from the total eleven evaluations. The sampling process has been repeated for 500 times and the corresponding power curves are drawn with transparent grey lines. Uneven distributions including normal and lognormal distributions produce lower power for both types of residual departures. Meanwhile, the lognormal distribution as an asymmetric distribution produce the lowest power.}\label{fig:different-x-dist-poly-power}
\end{figure}

@@

\hypertarget{acknowledgement}{%
\section{Acknowledgement}\label{acknowledgement}}

This research is enabled by the use of the following \texttt{R}
packages: cli, curl, dplyr, ggplot2, jsonlite, lmtest, mpoly, progress,
tibble, ggmosaic, purrr, tidyr, readr, stringr, here, kableExtra,
patchwork, rcartocolor. The study website is powered by pythonAnywhere
and the Flask web framework. The jsPsych framework is used to create
behavioral experiments that run in our study website.

The article was created with rticles (ref here), knitr (ref here) and R
Markdown (ref here). The project's Github repository (link here)
contains all materials required to reproduce this article.

\hypertarget{supplementary-material}{%
\section{Supplementary material}\label{supplementary-material}}

@@

\newpage

>>>>

\title{A Plot is Worth a Thousand Tests: Assessing Residual Diagnostics
with the Lineup Protocol}

@@

Figure \ref{fig:different-x-dist-poly-power} illustrates the impact of
fitted value distribution on the power of conventional tests and visual
tests. For conventional tests, only the power curves of appropriate
tests which are RESET tests and Breusch-Pagan tests for non-linearity
pattern and heteroskedasticity pattern respectively are shown. For
visual tests, note that we have collected more evaluations on lineup
with uniform fitted value distribution, and the number of evaluations on
a lineup will affect the power of the visual test. To have a fair
comparison, for lineups with uniform fitted value distribution, we
randomly sample five evaluations from the total eleven evaluations to
estimate the power curves. The sampling process is repeated 500 times to
produce an indication of the variation of the power.

In terms of visual tests, we do not observe significant difference
between the power curve of the uniform distribution and the power curve
of the discrete uniform distribution. Normal and lognormal distributions
produce lower power than uniform and discrete uniform distributions in
regards of residual departure patterns, but the gap is larger when the
residual plot contains heteroskedasticity pattern. This suggests that
residual departures displayed with evenly distributed fitted values are
visually more attractive and effective. Besides, the skewness of the
fitted value distribution plays a role in the degree of presence of
visual features as lognormal distribution produce the lowest power for
both non-linearity and heteroskedasticity patterns. A symmetric fitted
value distribution can better reveal the underlying shape.

Although it is unusual to consider fitted value distribution or
predictor distribution in power analysis, it is not so surprised that
the power of conventional tests vary because of fitted value
distribution. For instance, the power of the RESET test which is
effectively a F test depends on the realizations of the additional
predictors included in the auxiliary regression equation
\citep{jamshidian2007study}. And those additional predictors are linear
transformation of random vectors following certain predictor
distributions. What is truly unexpected is the huge variability in power
of conventional tests compared to visual tests. The discrete uniform
distribution constantly produce the highest power while the normal
distribution and the lognormal distribution produce the lowest power for
non-linearity and heteroskedasticity pattern respectively. The
difference in power corresponding to different distributions peak at
around 90\% when the effect size is at a moderate level. This huge
difference is undesirable as is not uncommon for analysts to omit the
factor of predictor distribution in power calculation and sample size
calculation. It also suggests that, under the designed scenarios in this
paper, the visual test is a more robust test against fitted value
distributions than a conventional test.

@@

\caption{Comparison of power on lineups with different fitted value distributions for conventional tests and visual tests. Power curves of conventional tests for non-linearity and heteroskedasticity patterns are produced by RESET tests and Breusch-Pagan tests respectively. Power curves of visual tests are estimated using five evaluations on each lineup. For lineups with uniform fitted value distribution, the five evaluations are randomly sampled from the total eleven evaluations. The sampling process has been repeated for 500 times and the corresponding power curves are drawn with transparent grey lines. The fitted value distribution has greater impact on the power of conventional tests than visual tests. Meanwhile, uneven distributions including normal and lognormal distributions genearlly produce lower power.}\label{fig:different-x-dist-poly-power}
\end{figure}

@@

\hypertarget{acknowledgements}{%
\section*{Acknowledgements}\label{acknowledgements}}
\addcontentsline{toc}{section}{Acknowledgements}

These \texttt{R} packages are used for the work: \texttt{cli}
\citep{cli}, \texttt{curl} \citep{curl}, \texttt{dplyr} \citep{dplyr},
\texttt{ggplot2} \citep{ggplot2}, \texttt{jsonlite} \citep{jsonlite},
\texttt{lmtest} \citep{lmtest}, \texttt{mpoly} \citep{mpoly},
\texttt{progress} \citep{progress}, \texttt{tibble} \citep{tibble},
\texttt{ggmosaic} \citep{ggmosaic}, \texttt{purrr} \citep{purrr},
\texttt{tidyr} \citep{tidyr}, \texttt{readr} \citep{readr},
\texttt{stringr} \citep{stringr}, \texttt{here} \citep{here},
\texttt{kableExtra} \citep{kableextra}, \texttt{patchwork}
\citep{patchwork}, \texttt{rcartocolor} \citep{rcartocolor}. The study
website is powered by \texttt{PythonAnywhere} \citep{pythonanywhere} and
the \texttt{Flask} web framework \citep{flask}. The \texttt{jsPsych}
framework \citep{jspsych} is used to create behavioral experiments that
run in our study website.

The article was created with R packages \texttt{rticles}
\citep{rticles}, \texttt{knitr} \citep{knitr} and \texttt{rmarkdown}
\citep{rmarkdown}. The project's Github repository (link here) contains
all materials required to reproduce this article.

\hypertarget{supplementary-material}{%
\section*{Supplementary material}\label{supplementary-material}}
\addcontentsline{toc}{section}{Supplementary material}

@@
